---
title: "Activity: Batch Processing"
output: html_notebook
---

## Setup

Run this chunk first. It loads all required packages, sets your API key, loads the essay data, and defines the `call_claude()` and `score_essay()` functions needed for this activity.

```{r setup}
library(httr)
library(jsonlite)
library(dplyr)
library(purrr)
library(glue)
library(DT)
library(lubridate)

if (Sys.getenv("ANTHROPIC_API_KEY") == "") {
  Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))
}

# Load student essay data
load(url("https://raw.githubusercontent.com/runyoncr/AIMECON_R_WORKSHOP/main/data/student_essays.Rdata"))

call_claude <- function(prompt,
                        model = "claude-sonnet-4-6",
                        system = NULL,
                        temperature = 0.5,
                        max_tokens = 4096,
                        effort = "low") {
  api_key <- Sys.getenv("ANTHROPIC_API_KEY")
  messages <- list(list(role = "user", content = prompt))
  request_body <- list(
    model = model,
    messages = messages,
    max_tokens = max_tokens,
    temperature = temperature,
    output_config = list(effort = effort)
  )
  if (!is.null(system)) {
    request_body$system <- system
  }
  headers <- add_headers(
    "x-api-key" = api_key,
    "anthropic-version" = "2023-06-01",
    "content-type" = "application/json"
  )
  response <- POST(
    url = "https://api.anthropic.com/v1/messages",
    headers,
    body = toJSON(request_body, auto_unbox = TRUE)
  )
  if (http_status(response)$category != "Success") {
    stop(paste("API request failed:", http_status(response)$message,
               "\nDetails:", content(response, "text", encoding = "UTF-8")))
  }
  result <- fromJSON(content(response, "text", encoding = "UTF-8"))
  return(result$content$text)
}

score_essay <- function(essay) {
  prompt <- glue(
    "You are an expert essay grader with experience in evaluating argumentative writing
    at the secondary and post-secondary level.

    ## TASK
    Score the student essay provided below using the three criteria in the rubric.
    Each criterion is scored from 1 to 5 in increments of 0.5 (1, 1.5, 2, 2.5, 3,
    3.5, 4, 4.5, 5). Apply each criterion independently before calculating the total.

    ## ESSAY PROMPT GIVEN TO STUDENT
    'Do you think that smartphones have destroyed communication among family and friends?
    Give specific reasons and details to support your opinion.'

    ## SCORING RUBRIC

    ### CRITERION 1: Content (1-5)
    Score the degree to which the essay develops a clear argument with relevant reasons
    and specific supporting examples in response to the essay prompt.

    - 5: Argument is thoroughly developed with multiple specific, relevant reasons and
         concrete examples. All content directly supports the central claim.
    - 4: Argument is well-developed with mostly specific reasons and examples. Minor
         gaps in development or relevance.
    - 3: Argument is adequately developed but reasons are sometimes general or examples
         are vague. Some content may be loosely connected to the argument.
    - 2: Argument is underdeveloped. Reasons are largely general or repetitive and
         examples are missing or unclear.
    - 1: Little to no recognizable argument. Content is largely irrelevant, missing,
         or does not respond to the prompt.

    Scores between these anchors (e.g., 1.5, 2.5) should be used when the essay falls
    between two descriptors.

    ### CRITERION 2: Organization (1-5)
    Score the degree to which the essay is logically structured at both the essay level
    (introduction, body, conclusion) and the paragraph level (topic sentences, coherence
    devices, single main idea per paragraph).

    - 5: Essay has a clear introduction with a thesis, well-organized body paragraphs
         each focused on a single idea, and a conclusion. Transitions and coherence
         devices are used effectively throughout.
    - 4: Essay structure is clear and mostly effective. Minor issues with transitions
         or paragraph focus.
    - 3: Basic structure is present but inconsistently applied. Some paragraphs may
         lack focus or transitions may be absent or mechanical.
    - 2: Structure is difficult to follow. Paragraphs may lack topic sentences or blend
         multiple unrelated ideas. Few or no transitions.
    - 1: No discernible organizational structure. Ideas are presented randomly with no
         paragraph logic.

    ### CRITERION 3: Language (1-5)
    Score the overall quality of language use across three equally weighted
    sub-dimensions: (a) vocabulary range and accuracy, (b) grammar and usage
    correctness, and (c) spelling and punctuation accuracy. Average across these three
    sub-dimensions to arrive at a single Language score.

    - 5: (a) Sophisticated and varied vocabulary with accurate collocations; (b) grammar
         and usage are correct throughout; (c) spelling and punctuation are correct
         throughout.
    - 4: (a) Good vocabulary range with occasional imprecision; (b) mostly correct
         grammar with minor errors that do not impede meaning; (c) few spelling or
         punctuation errors.
    - 3: (a) Adequate but limited vocabulary, some word choice errors; (b) grammar
         errors are noticeable but meaning is generally clear; (c) some spelling and
         punctuation errors.
    - 2: (a) Narrow or frequently inaccurate vocabulary; (b) grammar errors are frequent
         and sometimes obscure meaning; (c) spelling and punctuation errors are frequent.
    - 1: (a) Very limited vocabulary with pervasive errors; (b) grammar errors throughout
         severely impede meaning; (c) spelling and punctuation errors throughout.

    ## HANDLING EDGE CASES
    - If the essay is off-topic or does not respond to the prompt, assign a Content
      score of 1 and note this in the justification.
    - If the essay is fewer than 50 words, assign a maximum score of 2 for Content
      and Organization, and score Language based on what is present.

    ## STUDENT ESSAY TO SCORE
    {essay}

    ## OUTPUT INSTRUCTIONS
    Return your response ONLY as a valid JSON object with no additional text before or
    after it. Use the following structure exactly:

    {{
      'content_score': <number>,
      'content_justification': '<1-2 sentence justification>',
      'organization_score': <number>,
      'organization_justification': '<1-2 sentence justification>',
      'language_score': <number>,
      'language_justification': '<1-2 sentence justification referencing vocabulary,
                                  grammar, and mechanics>',
      'total_score': <sum of three scores>
    }}

    Ensure that 'total_score' equals the sum of the three criterion scores. Do not
    include any text outside the JSON object."
  )
  return(prompt)
}

# Build scoring prompts for all essays
student_essays$scoring_prompt <- sapply(student_essays$essay, score_essay)

# Score the first essay to confirm the prompt is working
student_essays$scoring_ouptut <- NA
student_essays$scoring_ouptut[1] <- call_claude(student_essays$scoring_prompt[1])
cat(student_essays$scoring_ouptut[1])
```

---

# Activity: Batch Processing

In the previous activity, we developed and refined a scoring prompt and tested it on a single essay. In the overview chapter, we saw that batch processing offers a path from that proof-of-concept to something operationally useful: consistent, traceable, and cost-efficient scoring at scale.

This activity works through that progression in two parts. First, we'll extend the single-essay call from the previous activity into a full loop-based scoring run — the approach you'd reach for when batch processing isn't available or isn't needed. Then we'll implement the same task using Anthropic's Batch API, compare the approaches, and use the batch infrastructure to explore a question that should feel natural to measurement professionals: how consistent is the model's scoring across repeated calls?

---

## Part 1: Scoring Essays in a Loop

### Parsing the JSON Output

Before we can loop across all essays, we need a reliable way to extract scores from the model's JSON response. The `score_essay()` prompt instructs the model to return a JSON object with six fields (three scores and three justifications) plus a `total_score`. The function below handles the extraction, including stripping any markdown fencing the model might add despite being asked not to.

```{r parse-scores}
parse_scores <- function(response_text) {
  clean <- gsub("```(?:json)?\\s*|```", "", response_text, perl = TRUE)
  clean <- trimws(clean)

  tryCatch({
    scores <- fromJSON(clean)
    data.frame(
      content      = scores$content_score,
      organization = scores$organization_score,
      language     = scores$language_score,
      total        = scores$total_score,
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    warning("Could not parse response: ", e$message)
    data.frame(content = NA_real_, organization = NA_real_,
               language = NA_real_, total = NA_real_)
  })
}
```

Verify it works on the essay already scored in setup:

```{r verify-parse}
parse_scores(student_essays$scoring_ouptut[1])
```

### Running the Full Loop

With `parse_scores()` in hand, we can score all essays using `purrr::map()`. We'll time the run so we have a concrete reference point for the batch comparison coming up.

```{r full-loop}
start_time <- Sys.time()

loop_results <- student_essays |>
  mutate(
    raw_response = map_chr(scoring_prompt, call_claude),
    scores       = map(raw_response, parse_scores)
  ) |>
  tidyr::unnest(scores)

end_time <- Sys.time()

cat("Essays scored:", nrow(loop_results), "\n")
cat("Time elapsed: ", round(difftime(end_time, start_time, units = "secs"), 1), "seconds\n")
```

<details>
<summary><strong>Note on sequential processing</strong></summary>

Because each `map_chr()` call waits for a response before moving to the next, the total runtime grows linearly with the number of essays. For a small dataset like this one it's perfectly fine, but for hundreds (or thousands!) of records it becomes the bottleneck that motivates everything in Part 2.

</details>

### Inspecting the Results

```{r inspect-loop}
loop_results |>
  select(id, content, organization, language, total) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  arrange(total) |>
  datatable(options = list(pageLength = 10),
            caption = "Loop-based scores — one call per essay")
```

---

## Part 2: Batch Processing

The loop above works, but it has two limitations worth addressing as you scale up. First, it's slow: requests are sequential, so the total time is the sum of every individual API round-trip. Second, it costs more: Anthropic's Batch API reduces both input and output token prices by 50%, so for identical work, a batch job costs half as much as the same number of sequential calls.

The trade-off is that batch jobs aren't instantaneous — they're queued and processed asynchronously, with most completing within an hour. For any scoring or evaluation task that doesn't require real-time results, that's an easy trade to make.

The five steps below mirror the structure described in the overview chapter: create the request objects, write them to JSONL, submit the job, poll for completion, and retrieve and parse the results.

### Step 1: Create Batch Requests

Each request in a batch is a JSON object that contains a `custom_id` (your identifier, returned with the result so you can match responses back to inputs) and a `params` block that mirrors the body of a standard `/v1/messages` call.

```{r create-batch-requests, eval = FALSE}
create_batch_requests <- function(essays_df) {
  lapply(seq_len(nrow(essays_df)), function(i) {
    list(
      custom_id = paste0("essay_", essays_df$id[i]),
      params = list(
        model      = "claude-sonnet-4-6",
        max_tokens = 1024,
        messages   = list(
          list(role = "user", content = essays_df$scoring_prompt[i])
        )
      )
    )
  })
}

batch_requests <- create_batch_requests(student_essays)

cat("Requests created:", length(batch_requests), "\n")
```

### Step 2: Write Requests to JSONL

The Batch API expects requests in JSONL format — one JSON object per line, no surrounding array. `toJSON()` with `auto_unbox = TRUE` handles the serialization; `writeLines()` handles the one-per-line format.

```{r write-batch-file, eval = FALSE}
write_batch_file <- function(batch_requests, output_file = "batch_requests.jsonl") {
  jsonl_lines <- sapply(batch_requests, function(req) {
    toJSON(req, auto_unbox = TRUE)
  })
  writeLines(jsonl_lines, output_file)
  cat("Written to:", output_file, "\n")
  invisible(output_file)
}

batch_jsonl <- write_batch_file(batch_requests)
```

### Step 3: Submit the Batch Job

Submitting sends the JSONL contents as the `requests` field of a POST to `/v1/messages/batches`. The response includes a batch ID that you'll use to check status and retrieve results.

```{r submit-batch, eval = FALSE}
submit_batch <- function(jsonl_file) {
  jsonl_lines   <- readLines(jsonl_file)
  requests_list <- lapply(jsonl_lines, fromJSON, simplifyVector = FALSE)

  response <- POST(
    url = "https://api.anthropic.com/v1/messages/batches",
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01",
      "content-type"      = "application/json"
    ),
    body   = toJSON(list(requests = requests_list), auto_unbox = TRUE),
    encode = "json"
  )

  if (status_code(response) != 200) {
    stop("Batch submission failed: ", content(response, "text"))
  }

  result <- content(response, "parsed")
  cat("Batch submitted successfully!\n")
  cat("Batch ID:", result$id, "\n")
  cat("Status:  ", result$processing_status, "\n")
  invisible(result)
}

batch_info <- submit_batch(batch_jsonl)
```

### Step 4: Check Batch Status

Batches process asynchronously. Poll this endpoint until `processing_status` is `"ended"`. For a small job like this one, it typically takes under two minutes.

```{r check-batch-status, eval = FALSE}
check_batch_status <- function(batch_id) {
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id),
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )

  if (status_code(response) != 200) {
    stop("Status check failed: ", content(response, "text"))
  }

  result <- content(response, "parsed")
  cat("Status:     ", result$processing_status, "\n")
  cat("Succeeded:  ", result$request_counts$succeeded, "\n")
  cat("Errored:    ", result$request_counts$errored, "\n")
  invisible(result)
}

batch_status <- check_batch_status(batch_info$id)
```

<details>
<summary><strong>Tip: Polling for completion</strong></summary>

For jobs you'll leave running while you move on to other work, you can wrap `check_batch_status()` in a polling loop that checks every 30 seconds and prints a message when the job finishes. For our workshop setting, we'll just call it manually a minute or two after submitting.

</details>

### Step 5: Retrieve and Parse Results

Once `processing_status` is `"ended"`, fetch the results. The response is JSONL again — one result object per line — so we split, parse, and extract the score from each.

```{r get-batch-results, eval = FALSE}
get_batch_results <- function(batch_id) {
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id, "/results"),
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )

  if (status_code(response) != 200) {
    stop("Result retrieval failed: ", content(response, "text"))
  }

  results_text  <- content(response, "text", encoding = "UTF-8")
  results_lines <- strsplit(results_text, "\n")[[1]]
  results_lines <- results_lines[nzchar(results_lines)]
  lapply(results_lines, fromJSON, simplifyVector = FALSE)
}

extract_batch_scores <- function(results_list) {
  rows <- lapply(results_list, function(result) {
    if (result$result$type != "succeeded") {
      warning("Request failed: ", result$custom_id, " — ", result$result$error$message)
      return(NULL)
    }
    response_text <- result$result$message$content[[1]]$text
    scores <- parse_scores(response_text)
    scores$essay_id <- sub("^essay_", "", result$custom_id)
    scores
  })
  do.call(rbind, Filter(Negate(is.null), rows))
}

batch_results <- get_batch_results(batch_info$id)
batch_scores  <- extract_batch_scores(batch_results)
```

### Comparing Loop and Batch Results

Since both runs used the same prompt and model, scores should be very similar. Differences, if any, reflect natural sampling variation at the default temperature.

```{r compare-results, eval = FALSE}
# How long did the batch take?
batch_start <- ymd_hms(batch_status$created_at)
batch_end   <- ymd_hms(batch_status$ended_at)
cat("Batch run time:", round(difftime(batch_end, batch_start, units = "mins"), 2), "minutes\n")

# Side-by-side score comparison
loop_scores <- loop_results |>
  select(id, content, organization, language, total) |>
  rename(essay_id = id)

comparison <- loop_scores |>
  mutate(essay_id = as.character(essay_id)) |>
  rename_with(~ paste0("loop_", .x), -essay_id) |>
  left_join(
    batch_scores |>
      mutate(essay_id = as.character(essay_id)) |>
      rename_with(~ paste0("batch_", .x), -essay_id),
    by = "essay_id"
  )

datatable(comparison |> mutate(across(where(is.numeric), ~ round(.x, 2))),
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Loop vs. Batch scores — same prompt, same model")
```

---

## Extension: Exploring Scoring Consistency

The loop and batch runs above each scored each essay once. That's appropriate for production scoring, but as measurement professionals you'll rightly ask: how stable are those scores? Does the model arrive at the same score if you ask again, and does that depend on how much randomness is in the sampling process?

The `temperature` parameter controls that randomness. At the default value in our function (0.5), the model samples probabilistically from its output distribution, which means repeated calls might produce different scores for the same essay. At `temperature = 0`, the model always selects the highest-probability token at each step — so repeated calls should produce nearly identical output.

Batch processing makes this kind of multi-rep study cheap and fast. The following code submits 20 calls per essay (200 requests total) at the default temperature and at `temperature = 0`, so we can see whether and how much scores vary.

```{r multi-rep-helpers, eval = FALSE}
# Helper: create n_reps requests per essay with a given temperature
create_multi_rep_requests <- function(essays_df, n_reps = 20, temperature = NULL) {
  requests <- list()
  k <- 1L

  for (i in seq_len(nrow(essays_df))) {
    params <- list(
      model      = "claude-sonnet-4-6",
      max_tokens = 1024,
      messages   = list(list(role = "user", content = essays_df$scoring_prompt[i]))
    )
    if (!is.null(temperature)) params$temperature <- temperature

    for (rep in seq_len(n_reps)) {
      requests[[k]] <- list(
        custom_id = paste0("essay_", essays_df$id[i], "_rep_", rep),
        params    = params
      )
      k <- k + 1L
    }
  }

  cat("Created", length(requests), "requests for",
      nrow(essays_df), "essays ×", n_reps, "repetitions\n")
  invisible(requests)
}

# Helper: extract scores from a multi-rep batch, preserving essay_id and rep
extract_multirep_scores <- function(results_list) {
  rows <- lapply(results_list, function(result) {
    if (result$result$type != "succeeded") {
      warning("Failed: ", result$custom_id)
      return(NULL)
    }
    response_text   <- result$result$message$content[[1]]$text
    scores          <- parse_scores(response_text)
    parts           <- strsplit(result$custom_id, "_")[[1]]
    scores$essay_id <- parts[2]
    scores$rep      <- as.integer(parts[4])
    scores
  })
  do.call(rbind, Filter(Negate(is.null), rows))
}
```

```{r default-temp-batch, eval = FALSE}
# Default temperature
default_temp_requests <- create_multi_rep_requests(student_essays, 
                                                   n_reps = 20)
default_temp_jsonl    <- write_batch_file(default_temp_requests, "batch_default_temp.jsonl")
default_temp_info     <- submit_batch(default_temp_jsonl)
```

```{r default-temp-retrieve, eval = FALSE}
# Check when finished, then retrieve
default_temp_status  <- check_batch_status(default_temp_info$id)
default_temp_results <- get_batch_results(default_temp_info$id)
default_temp_scores  <- extract_multirep_scores(default_temp_results)
```

```{r low-temp-batch, eval = FALSE}
# Temperature = 0
low_temp_requests <- create_multi_rep_requests(student_essays, 
                                               n_reps = 20, 
                                               temperature = 0)
low_temp_jsonl    <- write_batch_file(low_temp_requests, "batch_low_temp.jsonl")
low_temp_info     <- submit_batch(low_temp_jsonl)
```

```{r low-temp-retrieve, eval = FALSE}
low_temp_status  <- check_batch_status(low_temp_info$id)
low_temp_results <- get_batch_results(low_temp_info$id)
low_temp_scores  <- extract_multirep_scores(low_temp_results)
```

### Summarising Consistency

```{r summarise-scores, eval = FALSE}
summarise_scores <- function(df, label) {
  df |>
    group_by(essay_id) |>
    summarise(
      n             = n(),
      total_mean    = mean(total,        na.rm = TRUE),
      total_sd      = sd(total,          na.rm = TRUE),
      content_mean  = mean(content,      na.rm = TRUE),
      content_sd    = sd(content,        na.rm = TRUE),
      org_mean      = mean(organization, na.rm = TRUE),
      org_sd        = sd(organization,   na.rm = TRUE),
      lang_mean     = mean(language,     na.rm = TRUE),
      lang_sd       = sd(language,       na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(across(where(is.numeric), ~ round(.x, 2)),
           condition = label) |>
    arrange(total_mean)
}

default_summary  <- summarise_scores(default_temp_scores, "Default temperature")
low_temp_summary <- summarise_scores(low_temp_scores, "Temperature = 0")
```

```{r table-default-temp, eval = FALSE}
datatable(default_summary,
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Score consistency — default temperature (20 reps per essay)")
```

```{r table-low-temp, eval = FALSE}
datatable(low_temp_summary,
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Score consistency — temperature = 0 (20 reps per essay)")
```

<details>
<summary><strong>What to look for</strong></summary>

With default temperature you should see meaningful within-essay score variance (non-zero SDs), especially on the Language criterion where small differences in wording can tip a borderline response either way. At `temperature = 0` the SDs should collapse toward zero — or reach zero entirely — because the model is making the same deterministic choices on every call.

A few things worth reflecting on from a measurement standpoint:

- **Determinism vs. reliability.** Zero variance at `temperature = 0` might look like perfect reliability, but it's better understood as repeatability under identical conditions. A model that always assigns the same score to an essay hasn't necessarily converged on the *correct* score. And remember, even under conditions of `temperature = 0`, this does not _guarantee_ that the model will provide the same score every time.
- **Score distributions.** Compare the rank ordering of essays between the two conditions. If the rankings are stable but the variances differ, the temperature parameter is affecting precision rather than validity.
- **Cost implications.** These 400 requests (20 essays × 20 reps × 2 conditions) would have cost roughly twice as much without batch pricing. For a real inter-rater reliability study at scale, that 50% discount compounds quickly.

</details>
