# R Packages {#sec-r-packages}

```{r}
#| include: false
# This chunk activates knitr for inline code
```

R's strength lies in both its statistical capabilities and its open-source nature.
Its code is freely available to inspect, modify, and improve, supporting a global community of statisticians, data scientists, and developers.

R packages extend the language with reusable functions, data, and documentation.
Community members develop packages for many academic domains (see [CRAN Task Views](https://cran.r-project.org/web/views/){target="_blank"}), allowing you to build on established and emerging analytical methods.

## Installing packages {#workshop-package}

Start by installing the packages below from CRAN in your R console.
You generally install a package once per R library, then load it in each session where you need it.
Later activities introduce additional packages as needed.

```{r}
#| eval: false
install.packages(c("httr2", "jsonlite", "glue", "ellmer", "shiny"))
```

For the API connection examples, load these two packages:

```{r}
library(httr2)
library(jsonlite)
```

These commands normally complete without printing output.
If R reports that a package is missing, install it in the R session you are using, then try loading it again.

The workshop's [call_claude_AIMECON26.R helper](downloads/call_claude_AIMECON26.R) is an R script, not a package to install with `install.packages()`.
It defines `call_claude()` and uses `httr2` to communicate with Anthropic's API.

## Individual package details

Here are the main roles of these packages in the workshop:

| Package | Workshop use |
|---|---|
| [`httr2`](https://httr2.r-lib.org/) | Build and send HTTP requests to an API. |
| [`jsonlite`](https://cran.r-project.org/package=jsonlite) | Convert between R objects and JSON data. |
| [`glue`](https://glue.tidyverse.org/) | Insert values into reusable prompt templates. |
| [`ellmer`](https://ellmer.tidyverse.org/) | Manage conversations with LLMs from R. |
| [`shiny`](https://shiny.posit.co/r/getstarted/shiny-basics/lesson1/) | Build interactive web applications with R. |

The workshop helper uses `httr2`, while `jsonlite` is used in later examples for working directly with JSON data.

We'll explore prompt templates and JSON in [Practical Tools](11-practical-tools.qmd), and web applications in [R Shiny](25-shiny.qmd).

## Generative AI packages

Several R packages provide interfaces to large language models (LLMs).
We'll use [`ellmer`](https://ellmer.tidyverse.org/){target="_blank"} for [conversational interactions](13-conversations-via-api.qmd).
For other options, see the [LLM-specific R packages](30-reference-materials.qmd#sec-r-llm-packages) in the reference materials.

