# Testing API Connection {#sec-test-connect}

Many of the interactive portions of this workshop depend on successfully being able to connect to a generative AI model via API (we will discuss API keys in a _little_ more depth later).
For the purposes of this workshop I've set up an API key through Anthropic for you all to use (see @sec-workshop-key).

The following is a _very_ basic function for you to test that you can connect to [Anthropic's Claude Sonnet 4.5 model](https://www.anthropic.com/news/claude-sonnet-4-5){target="_blank"}. 
As we progress through the workshop we'll use a slightly different function that gives you more control over some of the model generation parameters.

To use this function you must have a .Renviron file in your working directory, as the function pulls the API key from that environment file.
I will securely distribute this file to workshop participants ([contact me](mailto:CRunyon@nbme.org) if you have not yet received this file).
If you've just added the .Renviron file to your directory, you'll need to refresh your R session in order for it to be registered in the environment.
You can do this via the tabs at the top of the R Studio interface: Session --> Restart R.

## `call_claude` function {#sec-call-claude}

You can download the following function by clicking the copy button (looks like a clipboard) in the code chunk displayed below.
Alternatively, you can click the icon below:

<a href="downloads/call_claude.R" download style="display: inline-block; padding: 10px 20px; background-color: #4A90E2; color: black; text-decoration: none; border-radius: 5px;">
  📥 Download `call_claude`
</a>

```{r call_claude}

# Required packages
library(httr)
library(jsonlite)

call_claude <- function(prompt,
                        model = "claude-sonnet-4-5-20250929") {
  
  # Get API key from environment
  # You will need to change this to your own API key after workshop
  api_key <- Sys.getenv("ANTHROPIC_API_KEY")

  # Convert text prompt to required message format
  messages <- list(list(role = "user", content = prompt))
  
  # Build request body
  request_body <- list(
    model = model,
    messages = messages,
    max_tokens = 1024 # Required; will be an argument in other functions
  )
  
  # Set up headers
  headers <- add_headers(
    "x-api-key" = api_key,
    "anthropic-version" = "2023-06-01",
    "content-type" = "application/json"
  )
  
  # Make the API request
  response <- POST(
    url = "https://api.anthropic.com/v1/messages",
    headers,
    body = toJSON(request_body, auto_unbox = TRUE)
  )
  
  # Check if request was successful
  if (http_status(response)$category != "Success") {
    stop(paste("API request failed:", http_status(response)$message, 
               "\nDetails:", content(response, "text", encoding = "UTF-8")))
  }
  
  # Parse response and extract text content
  result <- fromJSON(content(response, "text", encoding = "UTF-8"))
  return(as.character(result$content)[2])
}

```

The only argument that needs to be supplied to the function is the prompt that you want to send to the model. 
This should be a text string that is enclosed by parentheses: "Tell me a joke about educational measurement."

```{r}

test_joke <- call_claude("Tell me a joke about educational measurement.")

test_joke

```

A few things to note:

1. Generally speaking, the process used by a generative AI model involves predicting the next token to be selected and sampling from a distribution of possible options.
(We will discuss ways to control this later.)
A (sometimes, mostly) beautiful implication of this is that responses from generative AI models will often be different, even if the **exact** same prompt is used.
This is especially the case in the simple function above, as we haven't made an effort to tune the model generation parameters to achieve a deterministic or replicable result.
Thus, when _you_ call Anthropic and have it generate a joke for you using the same prompt, you will likely get different responses each time.
In fact, each time that I compile this book on GitHub the result changes!

2. You may notice that escape sequences (`\n`, `\t`, etc.) may be present because the information is returned in the JSON format.
One way to handle this is to wrap your response in `cat()`.
We'll revisit this again later in other sections.

```{r}

cat(test_joke)

```
