# Testing API Connection {#sec-test-connect}

Many of the interactive portions of this workshop depend on successfully being able to connect to a generative AI model via API (we will discuss API keys in more depth later). For this workshop you'll need to obtain your own Anthropic API key through [console.anthropic.com](https://console.anthropic.com). 

$5 in credit should be sufficient for all workshop purposes, and you'll have enough left over to continue practicing sending LLM calls via API.

The following is a basic function for testing your connection to an Anthropic Claude model. We'll cover some of its elements later. To use it, load your API key into the R session with the syntax below:

```{r api-key, eval = FALSE}
Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))
```

When using the API key in your workflow, you can skip this step by creating an .Renviron file that is saved in your working directory. Each API key should have a line like:
`ANTHROPIC_API_KEY='paste_your_api_key_here'`.

## `call_claude` function {#sec-call-claude}

You can download the following function by clicking the copy button (looks like a clipboard) in the code chunk displayed below.
Alternatively, you can click the icon below:

<a href="downloads/call_claude_AIMECON26.R" download style="display: inline-block; padding: 10px 20px; background-color: #4A90E2; color: black; text-decoration: none; border-radius: 5px;">
  📥 Download `call_claude_AIMECON26`
</a>

```{r call_claude}
source("downloads/call_claude_AIMECON26.R")


```

The only argument that _needs_ to be supplied to the function is the prompt that you want to send to the model; all other arguments have default values, and we'll discuss the other options in more detail in subsequent sections. 
The prompt should be a text string that is enclosed by parentheses: e.g., "Tell me a joke about educational measurement."

```{r}

test_joke <- call_claude("Tell me a joke about educational measurement.")

test_joke

```

A few things to note:

1. Generally speaking, the process used by a generative AI model involves predicting the next token to be selected and sampling from a distribution of possible options.
(We will discuss ways to better control this later.)
A (sometimes, mostly) beautiful implication of this is that responses from generative AI models will often be different, even if the **exact** same prompt is used.
This is especially the case in the simple function above, as we haven't made an effort to tune the [model generation parameters](08-gen-ai-parameters.qmd#sec-gen-params) to achieve a deterministic or replicable result.
Thus, when _you_ call Anthropic and have it generate a joke for you using the same prompt, you will likely get different responses each time.
In fact, each time that I compile this book on GitHub the result changes!

2. You may notice that escape sequences (`\n`, `\t`, etc.) may be present because the is returned from the model.
One way to handle this is to wrap your response in `cat()`.
We'll revisit this again later in other sections.

```{r}

cat(test_joke)

```
