# Testing API Connection {#sec-test-connect}

Many of the interactive portions of this workshop depend on successfully being able to connect to a generative AI model via API (we will discuss API keys in more depth later). For this workshop you'll need to obtain your own Anthropic API key through [console.anthropic.com](https://console.anthropic.com). 
$5 in credit should be sufficient for all workshop purposes, and you'll have enough left over to continue practicing what you've learned today.

The following is a _very_ basic function for you to test that you can connect to [Anthropic's Claude Sonnet 4.6 model](https://www.anthropic.com/news/claude-sonnet-4-6). 
As we progress through the workshop we'll use a slightly different function that gives you more control over some of the model generation parameters.

To use this function you'll need to have the API loaded for the R session, which can you do via the syntax below:

```{r api-key, eval = FALSE}
library(httr)
library(jsonlite)
Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))
```

When using the API key in your workflow, you can skip this step by creating an .Renviron file that is saved in your working directory. Each API key should have a line like:
`ANTHROPIC_API_KEY='paste_your_api_key_here'`.

## `call_claude` function {#sec-call-claude}

You can download the following function by clicking the copy button (looks like a clipboard) in the code chunk displayed below.
Alternatively, you can click the icon below:

<a href="downloads/call_claude_NCME26.R" download style="display: inline-block; padding: 10px 20px; background-color: #4A90E2; color: black; text-decoration: none; border-radius: 5px;">
  📥 Download `call_claude_NCME26`
</a>

```{r call_claude}

library(httr)
library(jsonlite)

call_claude <- function(prompt,
                        model = "claude-sonnet-4-6",
                        system = NULL,
                        temperature = 0.5,
                        max_tokens = 4096,
                        effort = "low") {
  
  # Get API key from environment
  api_key <- Sys.getenv("ANTHROPIC_API_KEY")
  
  # Convert text prompt to required message format
  messages <- list(list(role = "user", content = prompt))
  
  # Build request body
  request_body <- list(
    model = model,
    messages = messages,
    max_tokens = max_tokens,
    temperature = temperature,
    output_config = list(effort = effort)
  )
  
  # Add system prompt if provided
  if (!is.null(system)) {
    request_body$system <- system
  }
  
  # Set up headers
  headers <- add_headers(
    "x-api-key" = api_key,
    "anthropic-version" = "2023-06-01",
    "content-type" = "application/json"
  )
  
  # Make the API request
  response <- POST(
    url = "https://api.anthropic.com/v1/messages",
    headers,
    body = toJSON(request_body, auto_unbox = TRUE)
  )
  
  # Check if request was successful
  if (http_status(response)$category != "Success") {
    stop(paste("API request failed:", http_status(response)$message,
               "\nDetails:", content(response, "text", encoding = "UTF-8")))
  }
  
  # Parse response and extract text content
  result <- fromJSON(content(response, "text", encoding = "UTF-8"))
  return(result$content$text)
  
}


```

For now, the only argument that needs to be supplied to the function is the prompt that you want to send to the model; we'll discuss the other options in more detail in subsequent sections. 
The prompt should be a text string that is enclosed by parentheses: "Tell me a joke about educational measurement."

```{r}

test_joke <- call_claude("Tell me a joke about educational measurement.")

test_joke

```

A few things to note:

1. Generally speaking, the process used by a generative AI model involves predicting the next token to be selected and sampling from a distribution of possible options.
(We will discuss ways to better control this later.)
A (sometimes, mostly) beautiful implication of this is that responses from generative AI models will often be different, even if the **exact** same prompt is used.
This is especially the case in the simple function above, as we haven't made an effort to tune the [model generation parameters](08-gen-ai-parameters.qmd#sec-gen-params) to achieve a deterministic or replicable result.
Thus, when _you_ call Anthropic and have it generate a joke for you using the same prompt, you will likely get different responses each time.
In fact, each time that I compile this book on GitHub the result changes!

2. You may notice that escape sequences (`\n`, `\t`, etc.) may be present because the information is returned in the JSON format.
One way to handle this is to wrap your response in `cat()`.
We'll revisit this again later in other sections.

```{r}

cat(test_joke)

```
