# Activity: Batch Processing

Download the notebook for this activity:
<a href="downloads/activity-batch-processing.Rmd" download style="display: inline-block; padding: 10px 20px; background-color: #4A90E2; color: black; text-decoration: none; border-radius: 5px;">
  📥 Download `activity-batch-processing.Rmd`
</a>

In the previous activity, we developed and refined a scoring prompt and tested it on a single essay.
In the overview chapter, we saw that batch processing offers a path from that proof-of-concept to something operationally useful: consistent, traceable, and cost-efficient scoring at scale.

This activity works through that progression in two parts.
First, we'll extend the single-essay call from the previous activity into a full loop-based scoring run — the approach you'd reach for when batch processing isn't available or isn't needed.
Then we'll implement the same task using Anthropic's Batch API, compare the approaches, and use the batch infrastructure to explore a question that should feel natural to measurement professionals: how consistent is the model's scoring across repeated calls?

We'll pick up exactly where the previous activity left off: `student_essays` is loaded, `score_essay()` is defined, and we have a single scored essay in `student_essays$scoring_ouptut[1]` to confirm the prompt is working.

---

## Part 1: Scoring Essays in a Loop

### Parsing the JSON Output

Before we can loop across all essays, we need a reliable way to extract scores from the model's JSON response.
The `score_essay()` prompt instructs the model to return a JSON object with six fields (three scores and three justifications) plus a `total_score`.
The function below handles the extraction, including stripping any markdown fencing the model might add despite being asked not to.

```{r, eval = FALSE}

library(jsonlite)

parse_scores <- function(response_text) {
  # Strip markdown fencing if present (```json ... ```)
  clean <- gsub("```(?:json)?\\s*|```", "", response_text, perl = TRUE)
  clean <- trimws(clean)

  tryCatch({
    scores <- fromJSON(clean)
    data.frame(
      content      = scores$content_score,
      organization = scores$organization_score,
      language     = scores$language_score,
      total        = scores$total_score,
      stringsAsFactors = FALSE
    )
  }, error = function(e) {
    warning("Could not parse response: ", e$message)
    data.frame(content = NA_real_, organization = NA_real_,
               language = NA_real_, total = NA_real_)
  })
}

```

You can verify it works on the essay you already scored:

```{r, eval = FALSE}

parse_scores(student_essays$scoring_ouptut[1])

```

### Running the Full Loop

With `parse_scores()` in hand, we can score all essays using `purrr::map()`.
We'll time the run so we have a concrete reference point for the batch comparison coming up.

```{r, eval = FALSE}

library(dplyr)
library(purrr)

start_time <- Sys.time()

loop_results <- student_essays |>
  mutate(
    raw_response = map_chr(scoring_prompt, call_claude),
    scores       = map(raw_response, parse_scores)
  ) |>
  tidyr::unnest(scores)

end_time <- Sys.time()

cat("Essays scored:", nrow(loop_results), "\n")
cat("Time elapsed: ", round(difftime(end_time, start_time, units = "secs"), 1), "seconds\n")

```

::: {.callout-note}
Because each `map_chr()` call waits for a response before moving to the next, the total runtime grows linearly with the number of essays.
For a small dataset like this one it's perfectly fine, but for hundreds (or thousands!) of records it becomes the bottleneck that motivates everything in Part 2.
:::

### Inspecting the Results

```{r, eval = FALSE}

library(DT)

loop_results |>
  select(id, content, organization, language, total) |>
  mutate(across(where(is.numeric), ~ round(.x, 2))) |>
  arrange(total) |>
  datatable(options = list(pageLength = 10),
            caption = "Loop-based scores — one call per essay")

```

---

## Part 2: Batch Processing

The loop above works, but it has two limitations worth addressing as you scale up.
First, it's slow: requests are sequential, so the total time is the sum of every individual API round-trip.
Second, it costs more: Anthropic's Batch API reduces both input and output token prices by 50%, so for identical work, a batch job costs half as much as the same number of sequential calls.

The trade-off is that batch jobs aren't instantaneous — they're queued and processed asynchronously, with most completing within an hour.
For any scoring or evaluation task that doesn't require real-time results, that's an easy trade to make.

The five steps below mirror the structure described in the overview chapter: create the request objects, write them to JSONL, submit the job, poll for completion, and retrieve and parse the results.

### Step 1: Create Batch Requests

Each request in a batch is a JSON object that contains a `custom_id` (your identifier, returned with the result so you can match responses back to inputs) and a `params` block that mirrors the body of a standard `/v1/messages` call.

```{r, eval = FALSE}

create_batch_requests <- function(essays_df) {
  lapply(seq_len(nrow(essays_df)), function(i) {
    list(
      custom_id = paste0("essay_", essays_df$id[i]),
      params = list(
        model      = "claude-sonnet-4-20250514",
        max_tokens = 1024,
        messages   = list(
          list(role = "user", content = essays_df$scoring_prompt[i])
        )
      )
    )
  })
}

batch_requests <- create_batch_requests(student_essays)

cat("Requests created:", length(batch_requests), "\n")

```

### Step 2: Write Requests to JSONL

The Batch API expects requests in JSONL format — one JSON object per line, no surrounding array.
`toJSON()` with `auto_unbox = TRUE` handles the serialization; `writeLines()` handles the one-per-line format.

```{r, eval = FALSE}

write_batch_file <- function(batch_requests, output_file = "batch_requests.jsonl") {
  jsonl_lines <- sapply(batch_requests, function(req) {
    toJSON(req, auto_unbox = TRUE)
  })
  writeLines(jsonl_lines, output_file)
  cat("Written to:", output_file, "\n")
  invisible(output_file)
}

batch_jsonl <- write_batch_file(batch_requests)

```

### Step 3: Submit the Batch Job

Submitting sends the JSONL contents as the `requests` field of a POST to `/v1/messages/batches`.
The response includes a batch ID that you'll use to check status and retrieve results.

```{r, eval = FALSE}

library(httr)

submit_batch <- function(jsonl_file) {
  jsonl_lines    <- readLines(jsonl_file)
  requests_list  <- lapply(jsonl_lines, fromJSON, simplifyVector = FALSE)

  response <- POST(
    url = "https://api.anthropic.com/v1/messages/batches",
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01",
      "content-type"      = "application/json"
    ),
    body   = toJSON(list(requests = requests_list), auto_unbox = TRUE),
    encode = "json"
  )

  if (status_code(response) != 200) {
    stop("Batch submission failed: ", content(response, "text"))
  }

  result <- content(response, "parsed")
  cat("Batch submitted successfully!\n")
  cat("Batch ID:", result$id, "\n")
  cat("Status:  ", result$processing_status, "\n")
  invisible(result)
}

batch_info <- submit_batch(batch_jsonl)

```

### Step 4: Check Batch Status

Batches process asynchronously. Poll this endpoint until `processing_status` is `"ended"`.
For a small job like this one, it typically takes under two minutes.

```{r, eval = FALSE}

check_batch_status <- function(batch_id) {
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id),
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )

  if (status_code(response) != 200) {
    stop("Status check failed: ", content(response, "text"))
  }

  result <- content(response, "parsed")
  cat("Status:     ", result$processing_status, "\n")
  cat("Succeeded:  ", result$request_counts$succeeded, "\n")
  cat("Errored:    ", result$request_counts$errored, "\n")
  invisible(result)
}

batch_status <- check_batch_status(batch_info$id)

```

::: {.callout-tip}
For jobs you'll leave running while you move on to other work, you can wrap `check_batch_status()` in a polling loop that checks every 30 seconds and prints a message when the job finishes.
For our workshop setting, we'll just call it manually a minute or two after submitting.
:::

### Step 5: Retrieve and Parse Results

Once `processing_status` is `"ended"`, fetch the results.
The response is JSONL again — one result object per line — so we split, parse, and extract the score from each.

```{r, eval = FALSE}

get_batch_results <- function(batch_id) {
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id, "/results"),
    add_headers(
      "x-api-key"         = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )

  if (status_code(response) != 200) {
    stop("Result retrieval failed: ", content(response, "text"))
  }

  results_text  <- content(response, "text", encoding = "UTF-8")
  results_lines <- strsplit(results_text, "\n")[[1]]
  results_lines <- results_lines[nzchar(results_lines)]
  lapply(results_lines, fromJSON, simplifyVector = FALSE)
}


extract_batch_scores <- function(results_list) {
  rows <- lapply(results_list, function(result) {
    if (result$result$type != "succeeded") {
      warning("Request failed: ", result$custom_id, " — ", result$result$error$message)
      return(NULL)
    }

    # The content field is a list with one text block
    response_text <- result$result$message$content[[1]]$text

    scores <- parse_scores(response_text)
    scores$essay_id <- sub("^essay_", "", result$custom_id)
    scores
  })

  do.call(rbind, Filter(Negate(is.null), rows))
}


batch_results <- get_batch_results(batch_info$id)
batch_scores  <- extract_batch_scores(batch_results)

```

### Comparing Loop and Batch Results

Since both runs used the same prompt and model, scores should be very similar.
Differences, if any, reflect natural sampling variation at the default temperature.

```{r, eval = FALSE}

library(lubridate)

# How long did the batch take?
batch_start <- ymd_hms(batch_status$created_at)
batch_end   <- ymd_hms(batch_status$ended_at)
cat("Batch run time:", round(difftime(batch_end, batch_start, units = "mins"), 2), "minutes\n")

# Side-by-side score comparison
loop_scores <- loop_results |>
  select(id, content, organization, language, total) |>
  rename(essay_id = id)

comparison <- loop_scores |>
  mutate(essay_id = as.character(essay_id)) |>
  rename_with(~ paste0("loop_", .x), -essay_id) |>
  left_join(
    batch_scores |>
      mutate(essay_id = as.character(essay_id)) |>
      rename_with(~ paste0("batch_", .x), -essay_id),
    by = "essay_id"
  )

datatable(comparison |> mutate(across(where(is.numeric), ~ round(.x, 2))),
          options  = list(pageLength = 10, scrollX = TRUE),
          caption  = "Loop vs. Batch scores — same prompt, same model")

```

---

## Extension: Exploring Scoring Consistency

The loop and batch runs above each scored each essay once.
That's appropriate for production scoring, but as measurement professionals you'll rightly ask: how stable are those scores?
Does the model arrive at the same score if you ask again, and does that depend on how much randomness is in the sampling process?

The `temperature` parameter controls that randomness.
At the default value (~1.0), the model samples probabilistically from its output distribution, which means repeated calls can produce different scores for the same essay.
At `temperature = 0`, the model is effectively deterministic — it always selects the highest-probability token at each step — so repeated calls should produce identical output.

Batch processing makes this kind of multi-rep study cheap and fast.
The following code submits 20 calls per essay (200 requests total) at the default temperature and at `temperature = 0`, so we can see whether and how much scores vary.

```{r, eval = FALSE}

# Helper: create n_reps requests per essay with a given temperature
create_multi_rep_requests <- function(essays_df, n_reps = 20, temperature = NULL) {
  requests <- list()
  k <- 1L

  for (i in seq_len(nrow(essays_df))) {
    params <- list(
      model      = "claude-sonnet-4-20250514",
      max_tokens = 1024,
      messages   = list(list(role = "user", content = essays_df$scoring_prompt[i]))
    )
    if (!is.null(temperature)) params$temperature <- temperature

    for (rep in seq_len(n_reps)) {
      requests[[k]] <- list(
        custom_id = paste0("essay_", essays_df$id[i], "_rep_", rep),
        params    = params
      )
      k <- k + 1L
    }
  }

  cat("Created", length(requests), "requests for",
      nrow(essays_df), "essays ×", n_reps, "repetitions\n")
  invisible(requests)
}


# Helper: extract scores from a multi-rep batch, preserving essay_id and rep
extract_multirep_scores <- function(results_list) {
  rows <- lapply(results_list, function(result) {
    if (result$result$type != "succeeded") {
      warning("Failed: ", result$custom_id)
      return(NULL)
    }

    response_text <- result$result$message$content[[1]]$text
    scores        <- parse_scores(response_text)

    # custom_id format: "essay_<id>_rep_<rep>"
    parts           <- strsplit(result$custom_id, "_")[[1]]
    scores$essay_id  <- parts[2]
    scores$rep       <- as.integer(parts[4])
    scores
  })

  do.call(rbind, Filter(Negate(is.null), rows))
}

```

```{r, eval = FALSE}

# ── Default temperature ───────────────────────────────────────────────────────
default_temp_requests <- create_multi_rep_requests(student_essays, n_reps = 20)
default_temp_jsonl    <- write_batch_file(default_temp_requests, "batch_default_temp.jsonl")
default_temp_info     <- submit_batch(default_temp_jsonl)

```

```{r, eval = FALSE}

# Check when finished, then retrieve
default_temp_status  <- check_batch_status(default_temp_info$id)
default_temp_results <- get_batch_results(default_temp_info$id)
default_temp_scores  <- extract_multirep_scores(default_temp_results)

```

```{r, eval = FALSE}

# ── Temperature = 0 ───────────────────────────────────────────────────────────
low_temp_requests <- create_multi_rep_requests(student_essays, n_reps = 20, temperature = 0)
low_temp_jsonl    <- write_batch_file(low_temp_requests, "batch_low_temp.jsonl")
low_temp_info     <- submit_batch(low_temp_jsonl)

```

```{r, eval = FALSE}

low_temp_status  <- check_batch_status(low_temp_info$id)
low_temp_results <- get_batch_results(low_temp_info$id)
low_temp_scores  <- extract_multirep_scores(low_temp_results)

```

### Summarising Consistency

```{r, eval = FALSE}

summarise_scores <- function(df, label) {
  df |>
    group_by(essay_id) |>
    summarise(
      n             = n(),
      total_mean    = mean(total,        na.rm = TRUE),
      total_sd      = sd(total,          na.rm = TRUE),
      content_mean  = mean(content,      na.rm = TRUE),
      content_sd    = sd(content,        na.rm = TRUE),
      org_mean      = mean(organization, na.rm = TRUE),
      org_sd        = sd(organization,   na.rm = TRUE),
      lang_mean     = mean(language,     na.rm = TRUE),
      lang_sd       = sd(language,       na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(across(where(is.numeric), ~ round(.x, 2)),
           condition = label) |>
    arrange(total_mean)
}

default_summary <- summarise_scores(default_temp_scores, "Default temperature")
low_temp_summary <- summarise_scores(low_temp_scores, "Temperature = 0")

```

```{r, eval = FALSE}

datatable(default_summary,
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Score consistency — default temperature (20 reps per essay)")

```

```{r, eval = FALSE}

datatable(low_temp_summary,
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Score consistency — temperature = 0 (20 reps per essay)")

```

::: {.callout-note}
## What to look for

With default temperature you should see meaningful within-essay score variance (non-zero SDs), especially on the Language criterion where small differences in wording can tip a borderline response either way.
At `temperature = 0` the SDs should collapse toward zero — or reach zero entirely — because the model is making the same deterministic choices on every call.

A few things worth reflecting on from a measurement standpoint:

- **Determinism vs. reliability.** Zero variance at `temperature = 0` might look like perfect reliability, but it's better understood as repeatability under identical conditions. A model that always assigns the same score to an essay hasn't necessarily converged on the *correct* score. And remember, even under conditions of `temperature = 0`, this does not _guarantee_ that the model will provide the same score every time.
- **Score distributions.** Compare the rank ordering of essays between the two conditions. If the rankings are stable but the variances differ, the temperature parameter is affecting precision rather than validity.
- **Cost implications.** These 400 requests (20 essays × 20 reps × 2 conditions) would have cost roughly twice as much without batch pricing. For a real inter-rater reliability study at scale, that 50% discount compounds quickly.
:::
