# Suppressor Variables in Regression {#sec-example-statistical-suppression}

Can a variable with almost no correlation with an outcome still help predict it? A **suppressor variable** can strengthen another predictor's contribution by accounting for variation in that predictor that is unrelated to the outcome. Here we demonstrate this pattern using synthetic data and only base R.

## Generate Example Data

These are synthetic measurements for 120 students, not results from an actual study. We deliberately construct a `test_index` that mixes subject knowledge with an unrelated response-style component. The outcome, `score`, depends on knowledge plus random noise, but not on response style.

```{r}
#| label: ex-suppression-data
set.seed(2026)

knowledge <- rnorm(120)
response_style <- rnorm(120)

students <- data.frame(
  test_index = knowledge + response_style,
  response_style = response_style,
  score = 70 + 8 * knowledge + rnorm(120, mean = 0, sd = 3)
)

head(students)
```

Each row represents one student. `rnorm()` draws normally distributed values, and `set.seed()` makes the random draws reproducible. Knowledge and response style are generated independently, with mean 0 and standard deviation 1 in the population. Their sample correlation need not be exactly zero.

The `test_index` is an arbitrary numeric index, not a standardized test score. We retain response style as a measured predictor but leave `knowledge` out of the fitted models: the point is to see how the two observed predictors work together.

## Inspect the Correlations

`cor()` calculates the pairwise Pearson correlations. `round()` changes only their displayed precision.

```{r}
#| label: ex-suppression-correlations
correlations <- cor(students)
round(correlations, 2)
```

The test index correlates with the outcome (`r round(correlations["test_index", "score"], 2)`). Response style correlates with the test index (`r round(correlations["response_style", "test_index"], 2)`) but has little correlation with the outcome (`r round(correlations["response_style", "score"], 2)`). This is the pattern we built into the data: response style contributes to the index without contributing to the score-generation rule.

## Compare Two Regression Models

First predict scores using the test index alone. Then add response style. In an R formula, `~` separates the outcome from its predictors, and `+` includes both predictors without an interaction.

```{r}
#| label: ex-suppression-models
simple_model <- lm(score ~ test_index, data = students)
suppression_model <- lm(score ~ test_index + response_style, data = students)

round(coef(simple_model), 2)
round(coef(suppression_model), 2)
```

`lm()` fits each model, and `coef()` extracts its named coefficient estimates. Compare the `test_index` coefficients using the same students and the same measurement units:

```{r}
#| label: ex-suppression-comparison
simple_slope <- coef(simple_model)[["test_index"]]
adjusted_slope <- coef(suppression_model)[["test_index"]]
style_slope <- coef(suppression_model)[["response_style"]]

comparison <- data.frame(
  model = c("Test index only", "Test index + response style"),
  test_index_slope = c(simple_slope, adjusted_slope),
  r_squared = c(
    summary(simple_model)$r.squared,
    summary(suppression_model)$r.squared
  ),
  residual_sd = c(sigma(simple_model), sigma(suppression_model))
)

comparison[-1] <- round(comparison[-1], 2)
comparison
```

`summary()` supplies $R^2$, the proportion of sample outcome variation explained by the model. `sigma()` supplies the residual standard deviation, in score points. The `[["test_index"]]` expressions retrieve that predictor's coefficient by name; `comparison[-1]` selects the numeric columns for rounding.

| Quantity | Interpretation |
| --- | --- |
| Test-index slope, alone | One additional index unit is associated with `r round(simple_slope, 2)` additional predicted score points. |
| Test-index slope, adjusted | Holding response style constant, one additional index unit is associated with `r round(adjusted_slope, 2)` additional predicted score points. |
| Response-style slope, adjusted | Holding the test index constant, one additional response-style unit is associated with a `r round(style_slope, 2)`-point change in predicted score. |

Adding response style makes the positive test-index coefficient larger, increases $R^2$, and reduces the residual standard deviation. An increase in $R^2$ alone is not evidence of suppression: adding a predictor cannot decrease ordinary in-sample $R^2$. Here the coefficient change and correlation pattern are the important additional evidence.

## See the Coefficient Change

Both bars describe the same test-index unit, but they answer different questions: the first is an unadjusted association, and the second holds response style constant.

```{r}
#| label: ex-suppression-plot
#| fig-width: 6
#| fig-height: 4
#| fig-alt: "Two bars compare the test-index coefficient; the coefficient is larger after adjusting for response style."
barplot(
  c(simple_slope, adjusted_slope),
  names.arg = c("Index alone", "Adjusted for style"),
  col = c("steelblue4", "darkorange3"),
  ylim = c(0, max(simple_slope, adjusted_slope) * 1.15),
  ylab = "Score points per test-index unit",
  border = NA
)
abline(h = 0)
```

Why does this happen? In our construction, `test_index - response_style` equals `knowledge`. The second predictor lets the model account for the response-style component that obscures the knowledge-score association. In the population, the index-only slope is 4; including response style gives slopes of 8 for the index and -8 for response style. The fitted values differ because this is a finite sample with random noise.

The negative response-style coefficient does **not** mean response style causes lower scores. At a fixed test index, a larger response-style component implies less knowledge in this constructed example. Its conditional association can therefore be negative even when its marginal correlation with the outcome is near zero.

This illustrates **classical suppression**: the suppressor is uncorrelated with the outcome in the population but improves another predictor's contribution. Other forms of suppression need not have a zero marginal correlation. A variable's role depends on the other predictors and the outcome, so screening predictors solely by their individual outcome correlations can miss this pattern.

These are in-sample associations, not evidence of causality or guaranteed improvement on new students. The bars show point estimates; `confint(suppression_model)` gives coefficient confidence intervals under the model's assumptions. Real applications also require substantive justification for including a variable, model diagnostics, and assessment on new or held-out data.

For syntax, see the [R documentation for `lm()`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/lm.html). For the methodological distinction between suppression, mediation, and confounding, see [MacKinnon, Krull, and Lockwood (2000)](https://doi.org/10.1023/A:1026595011371).
