# Interactions in Linear Regression {#sec-example-regression-interactions}

Does an additional hour of studying have the same association with test scores for every study method? An **interaction** lets that slope differ between groups; here we fit and interpret one using only base R.

## Generate Example Data

These are synthetic scores for 120 students. We deliberately build in a steeper study-hours slope for the workshop group, then add random noise.

```{r}
#| label: ex-interaction-data
set.seed(2026)

students <- data.frame(
  hours = runif(120, min = 1, max = 9),
  method = factor(
    rep(c("Self-study", "Workshop"), each = 60),
    levels = c("Self-study", "Workshop")
  )
)

students$hours_centered <- students$hours - 5
workshop <- as.numeric(students$method == "Workshop")

students$score <- 65 +
  3 * students$hours_centered +
  4 * workshop +
  2 * students$hours_centered * workshop +
  rnorm(120, mean = 0, sd = 3)
```

With R's default treatment coding, the first factor level, `Self-study`, is the reference group. Subtracting five makes `hours_centered = 0` mean **five hours of studying**, not zero hours.

## Fit the Interaction

In an R model formula, `hours_centered * method` includes both predictors **and** their interaction: `hours_centered + method + hours_centered:method`. There is no need to create a product column yourself.

```{r}
#| label: ex-interaction-model
interaction_model <- lm(score ~ hours_centered * method, data = students)
coefficients <- coef(interaction_model)
round(coefficients, 2)
```

Read these estimates as follows:

| Coefficient | Interpretation |
| --- | --- |
| `(Intercept)` | Predicted self-study score at five hours: `r round(coefficients["(Intercept)"], 2)` points. |
| `hours_centered` | Self-study slope: `r round(coefficients["hours_centered"], 2)` additional predicted points per extra hour. |
| `methodWorkshop` | Workshop minus self-study difference **at five hours**: `r round(coefficients["methodWorkshop"], 2)` points. |
| `hours_centered:methodWorkshop` | Workshop minus self-study **slope difference**: `r round(coefficients["hours_centered:methodWorkshop"], 2)` points per hour. |

The interaction coefficient is **not** the workshop slope. Add it to the self-study slope:

```{r}
#| label: ex-interaction-slopes
self_study_slope <- coefficients[["hours_centered"]]
workshop_slope <- self_study_slope +
  coefficients[["hours_centered:methodWorkshop"]]

round(c("Self-study" = self_study_slope, "Workshop" = workshop_slope), 2)
```

An extra hour is associated with about `r round(self_study_slope, 2)` predicted points for self-study and `r round(workshop_slope, 2)` for workshop students. Neither coefficient is an overall slope averaged across both groups.

## Compare Predicted Scores

Use `predict()` to compare groups at specific hours. New data must use the **same centering value** as the training data.

```{r}
#| label: ex-interaction-predictions
predictions <- expand.grid(
  hours = c(2, 5, 8),
  method = levels(students$method)
)
predictions$hours_centered <- predictions$hours - 5
predictions$predicted_score <- predict(interaction_model, newdata = predictions)

data.frame(
  hours = predictions$hours,
  method = predictions$method,
  predicted_score = round(predictions$predicted_score, 1)
)
```

Compare rows with the same hours: the workshop minus self-study difference changes as hours increase. The `methodWorkshop` coefficient describes that difference only at five hours, not at every study duration.

```{r}
#| label: ex-interaction-plot
#| fig-width: 6
#| fig-height: 4
#| fig-alt: "Two fitted score lines from two to eight study hours; the workshop line is steeper than the self-study line."
self_study_predictions <- subset(predictions, method == "Self-study")
workshop_predictions <- subset(predictions, method == "Workshop")

plot(predicted_score ~ hours, data = self_study_predictions,
     type = "l", col = "steelblue4", lwd = 2,
     ylim = range(predictions$predicted_score),
     xlab = "Study hours", ylab = "Predicted score")
lines(predicted_score ~ hours, data = workshop_predictions,
      col = "darkorange3", lwd = 2, lty = 2)
legend("topleft", legend = c("Self-study", "Workshop"),
       col = c("steelblue4", "darkorange3"), lwd = 2, lty = c(1, 2), bty = "n")
```

The different slopes show the fitted interaction. These are point estimates, not guaranteed individual scores or evidence that either method causes improvement. For uncertainty, `confint(interaction_model)` gives coefficient confidence intervals; the interaction interval concerns the **difference in slopes**.

For syntax, see the [R documentation for `lm()`](https://stat.ethz.ch/R-manual/R-devel/library/stats/html/lm.html); for interpretation, see [UCLA's interaction guide](https://stats.oarc.ucla.edu/r/seminars/interactions-r/).