---
title: "Getting Started with the Anthropic API"
output: html_notebook
---

## Overview

In this workshop, we will be using the **Anthropic API** to interact with Claude directly from R. Before the workshop begins, this notebook will walk you through the two setup steps you need to complete:

1. Entering your Anthropic API key
2. Loading the `call_claude()` function

If both steps complete without errors, you are ready to go.

---

## Step 1: Load Required Packages

Run this chunk first. If either package is not installed, run `install.packages("httr")` or `install.packages("jsonlite")` in your console first, then try again.

```{r packages}
library(httr)
library(jsonlite)
```

---

## Step 2: Enter Your API Key

When you run the chunk below, a popup window will appear asking for your Anthropic API key. Paste your key into the box and click OK. Your key will be stored securely for this R session only — it is never saved to disk.

If you do not yet have an API key, you can create one at [console.anthropic.com](https://console.anthropic.com) under **API Keys**. You won't need many credits for the workshop - $5 should suffice.

```{r api-key}
Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))
```

---

## Step 3: Load the `call_claude()` Function

Run the chunk below to load the function we will be using throughout the workshop. You do not need to modify anything here.

```{r call-claude-function}
call_claude <- function(prompt,
                        model = "claude-sonnet-4-6",
                        system = NULL,
                        temperature = 0.5,
                        max_tokens = 4096,
                        effort = "low") {
  
  # Get API key from environment
  api_key <- Sys.getenv("ANTHROPIC_API_KEY")
  
  # Convert text prompt to required message format
  messages <- list(list(role = "user", content = prompt))
  
  # Build request body
  request_body <- list(
    model = model,
    messages = messages,
    max_tokens = max_tokens,
    temperature = temperature,
    output_config = list(effort = effort)
  )
  
  # Add system prompt if provided
  if (!is.null(system)) {
    request_body$system <- system
  }
  
  # Set up headers
  headers <- add_headers(
    "x-api-key" = api_key,
    "anthropic-version" = "2023-06-01",
    "content-type" = "application/json"
  )
  
  # Make the API request
  response <- POST(
    url = "https://api.anthropic.com/v1/messages",
    headers,
    body = toJSON(request_body, auto_unbox = TRUE)
  )
  
  # Check if request was successful
  if (http_status(response)$category != "Success") {
    stop(paste("API request failed:", http_status(response)$message,
               "\nDetails:", content(response, "text", encoding = "UTF-8")))
  }
  
  # Parse response and extract text content
  result <- fromJSON(content(response, "text", encoding = "UTF-8"))
  return(result$content$text)
  
}
```

---

## Step 4: Test Your Setup

Run the chunk below to send a simple message to Claude. If everything is working correctly, you should see a response appear below the chunk within a few seconds.

```{r test}
call_claude("Please respond with the following message exactly: 'API key confirmed. Ready for the workshop!'")
```

---

## You're All Set!

If you saw a response above, your API key is working and you are ready for the workshop. If you encountered an error, the most common causes are:

- The API key was entered incorrectly — re-run the Step 2 chunk and try again
- The `httr` or `jsonlite` package is not installed — see Step 1 above
- Your API key does not have available credits — check your usage at [console.anthropic.com](https://console.anthropic.com)
