# Activity: Conversations via API

Now it's time to practice a conversational interaction with an API.
We won't be spending a lot of time on this, but I would be remiss if you didn't at least get to kick the tires on this functionality in the workshop.

You can download the interactive notebook for this activity here:
<a href="downloads/activity-conversations-via-api.Rmd" download style="display: inline-block; padding: 10px 20px; background-color: #4A90E2; color: black; text-decoration: none; border-radius: 5px;">
  📥 Download `activity-conversations-via-api`
</a>

First you'll need to load the `ellmer` package and specify that we'll be using Claude Sonnet 4.6 for this activity:

```{r load ellmer, eval = FALSE}
library(ellmer)
```

Also load your API key if you haven't already provided it yet:

```{r load api key, eval = FALSE}
Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))
```

And let's specify that we'll be using Claude Sonnet 4.6 for this activity:

```{r start claude, eval = FALSE}
chat <- chat_anthropic(
  system_prompt = "You explain things concisely, focusing on only the most significant parts of the response.",
  model = "claude-sonnet-4-6")
```

Now that our chat is loaded, let's start!

## Console-based interaction

**Directions**: Have a short conversation with Claude in the console. It can be about anything; here are some suggestions for prompts.
- Idea 1
- Idea 2
- Idea 3

After a few turns, end the chat by submitting "Q" to quit.

```{r console, eval = FALSE}
live_console(chat)
```

## Browser-based interaction

Let's now see how the functionality changes for the browser-based version. You can use the same prompt, or start another one. Again, after a few turns, end the chat. 

```{r browser, eval = FALSE}
library(shiny)
library(shinychat)

chat$set_turns(list()) # Resets the conversation

live_browser(chat)
```

## Turn-based interaction

Lastly, let's try the final version of the chat where you send your prompts individually via R commands.

```{r turn 1, eval = FALSE}
chat$set_turns(list()) # Resets the conversation

# Paste your first prompt below
chat$chat(" ")
```

Next prompt:
```{r turn 2, eval = FALSE}
# Paste your second prompt below
chat$chat(" ")
```

And a third, final prompt:
```{r turn 3, eval = FALSE}
# Paste your third prompt below
chat$chat(" ")
```

Now review the turns in the conversation:
```{r rev turns, eval = FALSE}
turn_review <- chat$get_turns()

turn_review
```

## Silly system prompt
Now try a single interaction using a whimsical system prompt.

```{r silly system, eval = FALSE}
chat$set_turns(list()) # Resets the conversation

chat <- chat_anthropic(
  system_prompt = "You are an assistant that likes to respond in rhymes."
)

chat$chat("Briefly tell me the point of using a Rasch model.")
```