---
title: "Practical Tools: glue and JSON Formatting"
output: html_notebook
---

## Setup

Run this chunk first to load required packages and initialize your API connection.

```{r setup}
library(httr)
library(jsonlite)
library(glue)
library(purrr)

Sys.setenv(ANTHROPIC_API_KEY = rstudioapi::askForSecret("Anthropic API Key"))

call_claude <- function(prompt,
                        model = "claude-sonnet-4-6",
                        system = NULL,
                        temperature = 0.5,
                        max_tokens = 4096,
                        effort = "low") {
  api_key <- Sys.getenv("ANTHROPIC_API_KEY")
  messages <- list(list(role = "user", content = prompt))
  request_body <- list(
    model = model,
    messages = messages,
    max_tokens = max_tokens,
    temperature = temperature,
    output_config = list(effort = effort)
  )
  if (!is.null(system)) {
    request_body$system <- system
  }
  headers <- add_headers(
    "x-api-key" = api_key,
    "anthropic-version" = "2023-06-01",
    "content-type" = "application/json"
  )
  response <- POST(
    url = "https://api.anthropic.com/v1/messages",
    headers,
    body = toJSON(request_body, auto_unbox = TRUE)
  )
  if (http_status(response)$category != "Success") {
    stop(paste("API request failed:", http_status(response)$message,
               "\nDetails:", content(response, "text", encoding = "UTF-8")))
  }
  result <- fromJSON(content(response, "text", encoding = "UTF-8"))
  return(result$content$text)
}
```

---

# Practical Tools

The following are two practical tools that I have found helpful when using LLMs via API at scale.

## `glue`

The `glue` package is part of the [`tidyverse`](https://glue.tidyverse.org/articles/glue.html) and provides interpreted string literals for combining strings with data (string interpolation) in a readable and efficient way. It allows R expressions to be embedded directly within curly braces `{}` inside a string. This is a much safer and more efficient way of building a general prompt that will be used repeatedly with slightly different content, such as for content generation (different source material) or automated scoring (different rubrics and material to be evaluated).

Here is an example of what I mean.

```{r sample-variables}
# Specifying the input variables
sample_gender    <- "cisgender male"
sample_age       <- "37 years old"
sample_diagnosis <- "acute bronchitis"
```

Rather than building the prompt with `paste0()` like this:

```{r paste-example, eval = FALSE}
prompt <- paste0("Generate a multiple-choice question for a preclerkship medical student about a ",
                 age, " ", gender, " patient presenting with ", diagnosis, ".")
```

The `glue` package allows you to embed variables directly into a string using `{}`, making the prompt much easier to read, edit, and maintain. Run the chunk below to define the prompt function and generate a question for the sample patient:

```{r glue-prompt}
med_mcq_prompt <- function(gender, age, diagnosis) {
  glue("
    A {age} {gender} presents to the clinic with a new diagnosis of {diagnosis}.
    
    Generate a single best-answer multiple-choice question appropriate for a preclerkship 
    medical student that tests clinical reasoning about this presentation.
    
    The question should:
    - Be written in a clinical vignette format
    - Focus on diagnosis, pathophysiology, or first-line management
    - Include exactly 5 answer options (A through E)
    - Have one clearly correct answer and four plausible distractors
    - Do not include an explanation in your response. Include only the elements listed above.

    Format your response as follows:
    Question: [vignette and question stem]
    A. [option]
    B. [option]
    C. [option]
    D. [option]
    E. [option]
    Correct Answer: [letter]
  ")
}

one_med_mcq <- med_mcq_prompt(sample_gender, sample_age, sample_diagnosis)

cat(call_claude(one_med_mcq))
```

Notice that changing the patient's characteristics requires only updating the variables at the top — the prompt structure stays identical. This becomes especially powerful when iterating over many cases, allowing you to easily create an array of prompts to be submitted to an LLM without much fuss.

```{r prompt-array, eval = FALSE}
gender_array    <- c("cisgender female", "cisgender male", "cisgender male", "cisgender female")
age_array       <- c("23 years old", "37 years old", "63 years old", "52 years old")
diagnosis_array <- c("hyperthyroidism", "acute bronchitis", "atrial fibrillation", "hypertension")

med_prompt_array <- purrr::pmap_chr(
  list(gender_array, age_array, diagnosis_array),
  function(gender, age, diagnosis) med_mcq_prompt(gender, age, diagnosis)
)
```

---

# `JSON` Formatting

It is helpful to provide the model with an expected output structure. For the sake of demonstration, let's say that the output from the prompt above was sufficient for your assessment purpose. Best practices for storing assessment content are probably _not_ to just have the assessment item with the stem, response options, and answer key all stored together in a single text file. You'd likely want to extract these components[^1] and store them separately in your database. If you had the output structured as above, this would require the use of some of the functions in the [`stringr` package](https://stringr.tidyverse.org/), or, even worse — regular expressions! 😬

JSON is especially useful because it gives the model a clear, standardized structure to follow when generating output. It is a common data format used widely in APIs and software workflows, so models are generally quite good at producing it when asked. This makes the results much easier to parse, validate, and store automatically when working with large numbers of prompts. The [`jsonlite` package](https://cran.r-project.org/web/packages/jsonlite/refman/jsonlite.html) "offers simple, flexible tools for working with JSON in R, and is particularly powerful for building pipelines and interacting with a web API" — exactly our use case!

Let's revisit the above prompt and specify that the output should be in JSON format, with specific sections for the item stem, each response option, and the correct response:

```{r json-prompt}
med_mcq_prompt_json <- function(gender, age, diagnosis) {
  glue("
    A {age} {gender} presents to the clinic with a new diagnosis of {diagnosis}.
    
    Generate a single best-answer multiple-choice question appropriate for a preclerkship 
    medical student that tests clinical reasoning about this presentation.
    
    The question should:
    - Be written in a clinical vignette format
    - Focus on diagnosis, pathophysiology, or first-line management
    - Include exactly 5 answer options (A through E)
    - Have one clearly correct answer and four plausible distractors
    
    Return your response as a JSON object with the following structure:
    {{
      \"stem\": \"[vignette and question stem]\",
      \"options\": {{
        \"A\": \"[option]\",
        \"B\": \"[option]\",
        \"C\": \"[option]\",
        \"D\": \"[option]\",
        \"E\": \"[option]\"
      }},
      \"correct_answer\": \"[letter]\"
    }}
    
    Return only the JSON object. Do not include any additional text, explanation, or markdown formatting.
  ")
}

one_med_mcq_json    <- med_mcq_prompt_json(sample_gender, sample_age, sample_diagnosis)
one_med_json_output <- call_claude(one_med_mcq_json)

cat(one_med_json_output)
```

We can now use the `jsonlite` package to parse the JSON output into an R object:

```{r json-parse}
# Parse the JSON output from the model
result_parsed <- fromJSON(one_med_json_output)

# View the full structured output
result_parsed
```

The syntax below shows how you can extract additional elements from the JSON output. I have not shown all of the output here, but encourage you to go through the syntax below so you can become familiar with extracting specific elements from the parsed result.

```{r json-extract}
# Extract individual elements
result_parsed$stem
result_parsed$options
result_parsed$correct_answer

# Extract a specific response option
result_parsed$options$C

# Convert options to a named character vector (useful for downstream processing)
unlist(result_parsed$options)

# Combine into a tidy data frame
data.frame(
  option    = names(result_parsed$options),
  text      = unlist(result_parsed$options),
  correct   = names(result_parsed$options) == result_parsed$correct_answer,
  row.names = NULL
)
```

## `glue` and JSON Formatting

The `med_mcq_prompt_json` function above also highlights that the `glue` package can easily handle special characters and formatting when building prompts. It does require some extra steps (e.g., using double brackets `{{` `}}` instead of single brackets, the need for escape `\` for quotes and other formatting), but it is still _much_ easier than trying to do this with `paste`.

And, to be transparent, I often will ask an LLM to help me ensure that my function has the correct formatting for the JSON output — I don't know all of the minute details that always need to be included.

[^1]: Along with item meta-data, etc.
