# Activity: Batch Processing

### Batch Scoring

Now that we're repeating the same task across a batch of essays, it's a great time to introduce batch scoring!
This generally works as you'd think - you submit multiple tasks at once to the API to complete in parallel - with the one wrinkle that Anthropic requires that the requests be in the JSONL format.
In addition to being able to submit multiple requests at once, it's also cheaper to use than single-call API interactions.
As of this writing (Oct 19, 2025), you save 50% with batch pricing. 
This drops input costs from \$3 per 1M tokens to \$1.50 per 1M token, and output costs from \$15 per 1M tokens to \$7.50 per 1M tokens.

#### Creating Batch Requests

```{r, eval = FALSE}

# Step 1: Read essays and create batch requests
create_batch_requests <- function(essays_df, n_repetitions = 20){

  # Create a list to store batch requests
  batch_requests <- list()
  request_counter <- 1
  
  # Create n_repetitions requests for each essay (grouped by essay)
  for(i in 1:nrow(essays_df)){
    essay_prompt <- claude_all_rubric_prompt(essays_df$prompt[i], essays_df$essay[i])
    
    # Create 20 requests for this essay
    for(rep in 1:n_repetitions){
      request <- list(
        custom_id = paste0("essay_", essays_df$id[i], "_rep_", rep),
        params = list(
          model = "claude-sonnet-4-20250514",
          max_tokens = 1024,
          messages = list(
            list(
              role = "user",
              content = essay_prompt
            )
          )
        )
      )
      
      batch_requests[[request_counter]] <- request
      request_counter <- request_counter + 1
    }
  }
  
  cat("Created", length(batch_requests), "batch requests for", nrow(essays_df), "essays\n")
  
  return(batch_requests)
}

essay_batch_requests <- create_batch_requests(student_essays)

```

#### Writing Requests to JSONL

```{r, eval = FALSE}

# Step 2: Write requests to JSONL file
write_batch_file <- function(batch_requests, output_file = "batch_requests.jsonl"){
  # Write each request as a JSON line
  jsonl_lines <- sapply(batch_requests, function(req){
    toJSON(req, auto_unbox = TRUE)
  })
  
  writeLines(jsonl_lines, output_file)
  return(output_file)
}

essay_batch_jsonl <- write_batch_file(essay_batch_requests)

```

#### Submitting Batch Request

```{r, eval = FALSE}

# Step 3: Submit batch job to Anthropic
submit_batch <- function(jsonl_file){
  # Read and parse each line of the JSONL file
  jsonl_lines <- readLines(jsonl_file)
  
  # Parse each JSON line into a list
  requests_list <- lapply(jsonl_lines, function(line){
    fromJSON(line, simplifyVector = FALSE)
  })
  
  # Create batch
  response <- POST(
    url = "https://api.anthropic.com/v1/messages/batches",
    add_headers(
      "x-api-key" = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01",
      "content-type" = "application/json"
    ),
    body = toJSON(list(
      requests = requests_list
    ), auto_unbox = TRUE),
    encode = "json"
  )
  
  if(status_code(response) != 200){
    stop("Error submitting batch: ", content(response, "text"))
  }
  
  result <- content(response, "parsed")
  cat("Batch submitted successfully!\n")
  cat("Batch ID:", result$id, "\n")
  
  return(result)
}

essay_batch_info <- submit_batch(essay_batch_jsonl)

# Batch submitted successfully!
# Batch ID: msgbatch_018m5hGS9Lm3Gi28UorfsVTS

```

#### Checking Batch Status

```{r, eval = FALSE}

# Step 4: Check batch status
check_batch_status <- function(batch_id){
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id),
    add_headers(
      "x-api-key" = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )
  
  if(status_code(response) != 200){
    stop("Error checking batch status: ", content(response, "text"))
  }
  
  result <- content(response, "parsed")
  return(result)
}

essay_status <- check_batch_status(essay_batch_info$id)
# Finished!

library(lubridate)
batch_start_time <- ymd_hms(essay_status$created_at)
batch_stop_time <- ymd_hms(essay_status$ended_at)
batch_run_time <- as.numeric(batch_stop_time - batch_start_time)
# Time difference of 1.644671 mins
# Less than 2 minutes to score 200 essays!

```

#### Saving Batch Results

```{r, eval = FALSE}

# Step 5: Saving Batch Results

get_batch_results <- function(batch_id){
  response <- GET(
    url = paste0("https://api.anthropic.com/v1/messages/batches/", batch_id, "/results"),
    add_headers(
      "x-api-key" = Sys.getenv("ANTHROPIC_API_KEY"),
      "anthropic-version" = "2023-06-01"
    )
  )
  
  if(status_code(response) != 200){
    stop("Error retrieving results: ", content(response, "text"))
  }
  
  # Parse JSONL results
  results_text <- content(response, "text", encoding = "UTF-8")
  results_lines <- strsplit(results_text, "\n")[[1]]
  results_list <- lapply(results_lines[results_lines != ""], fromJSON)
  
  return(results_list)
}

essay_batch_results <- get_batch_results(essay_batch_info$id)
save(essay_batch_results, file = "data/essay_batch_results.Rdata")

```

#### Extracting Score from Results

```{r, eval = FALSE}

# Step 6: Extract scores from batch results
extract_batch_scores <- function(results_list, debug = FALSE){
  scores_df <- data.frame(
    essay_id = character(),
    repetition = integer(),
    content = numeric(),
    organization = numeric(),
    language = numeric(),
    total = numeric(),
    stringsAsFactors = FALSE
  )
  
  for(i in seq_along(results_list)){
    result <- results_list[[i]]
    
    if(result$result$type == "succeeded"){
      # Extract the response text - handle different possible structures
      message_content <- result$result$message$content
      
      # Debug: Print structure if requested
      if(debug && i == 1){
        cat("=== DEBUG: First result structure ===\n")
        cat("Content class:", class(message_content), "\n")
        cat("Content length:", length(message_content), "\n")
        if(is.list(message_content) && length(message_content) > 0){
          cat("First element class:", class(message_content[[1]]), "\n")
          print(str(message_content[[1]]))
        }
      }
      
      # Extract text from various possible structures
      response_text <- NULL
      
      if(is.data.frame(message_content)){
        # If it's a data frame, look for a 'text' column
        if("text" %in% names(message_content)){
          response_text <- message_content$text[1]
        }
      } else if(is.list(message_content) && length(message_content) > 0){
        first_elem <- message_content[[1]]
        
        if(is.list(first_elem)){
          # Check for $text field
          if(!is.null(first_elem$text)){
            response_text <- first_elem$text
          } else if(!is.null(first_elem$type) && first_elem$type == "text" && !is.null(first_elem$text)){
            response_text <- first_elem$text
          }
        } else if(is.character(first_elem)){
          response_text <- first_elem
        }
      } else if(is.character(message_content)){
        response_text <- message_content
      }
      
      # If we still don't have text, try to extract it more aggressively
      if(is.null(response_text) || length(response_text) == 0){
        # Try to find any 'text' field in the structure
        if(is.list(message_content)){
          text_fields <- unlist(lapply(message_content, function(x) {
            if(is.list(x) && "text" %in% names(x)) x$text else NULL
          }))
          if(length(text_fields) > 0) response_text <- text_fields[1]
        }
      }
      
      if(is.null(response_text) || length(response_text) == 0){
        warning("Could not extract text for ", result$custom_id)
        if(debug){
          cat("=== Full structure for ", result$custom_id, " ===\n")
          print(str(message_content))
        }
        next
      }
      
      # Debug: Print first response text
      if(debug && i == 1){
        cat("\n=== First response text (first 500 chars) ===\n")
        cat(substr(response_text, 1, 500), "\n\n")
      }
      
      # Extract scores using our previous function
      scores <- extract_scores(response_text, debug = debug && i == 1)
      
      # Debug: Check if scores were extracted
      if(debug && i == 1){
        cat("=== Extracted scores ===\n")
        print(scores)
      }
      
      # Parse custom_id to extract essay_id and repetition
      # Format: essay_2001_rep_1
      custom_id_parts <- strsplit(result$custom_id, "_")[[1]]
      essay_id <- custom_id_parts[2]
      repetition <- as.integer(custom_id_parts[4])
      
      scores_df <- rbind(scores_df, data.frame(
        essay_id = essay_id,
        repetition = repetition,
        content = scores["content"],
        organization = scores["organization"],
        language = scores["language"],
        total = scores["total"]
      ))
    } else {
      warning("Request ", result$custom_id, " failed: ", result$result$error$message)
    }
  }
  
  return(scores_df)
}


essay_batch_df <- extract_batch_scores(essay_batch_results, debug = TRUE)
```
```{r, eval = FALSE, echo = FALSE}
save(essay_batch_df, file = 'data/essay_batch_df.Rdata')
```
```{r, echo = FALSE}
load('data/essay_batch_df.Rdata')
```
```{r}
library(dplyr)
library(DT)

essay_summary_detailed <- essay_batch_df %>%
  group_by(essay_id) %>%
  summarise(
    n = n(),
    # Content
    content_mean = mean(content, na.rm = TRUE),
    content_sd = sd(content, na.rm = TRUE),
    content_min = min(content, na.rm = TRUE),
    content_max = max(content, na.rm = TRUE),
    # Organization
    org_mean = mean(organization, na.rm = TRUE),
    org_sd = sd(organization, na.rm = TRUE),
    org_min = min(organization, na.rm = TRUE),
    org_max = max(organization, na.rm = TRUE),
    # Language
    lang_mean = mean(language, na.rm = TRUE),
    lang_sd = sd(language, na.rm = TRUE),
    lang_min = min(language, na.rm = TRUE),
    lang_max = max(language, na.rm = TRUE),
    # Total
    total_mean = mean(total, na.rm = TRUE),
    total_sd = sd(total, na.rm = TRUE),
    total_min = min(total, na.rm = TRUE),
    total_max = max(total, na.rm = TRUE),
  ) %>%
  arrange(total_mean)

essay_summary_detailed <- essay_summary_detailed %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

datatable(essay_summary_detailed, 
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Summary Statistics by Essay (20 repetitions each)")


```

So we did see some variation, but I also used the default temperature. 

### Batch Scoring (`Temp = 0`)

Let's redo the batch call with `temperature = 0`. 

```{r, eval = FALSE}

create_lowtemp_requests <- function(essays_df, n_repetitions = 20){

  # Create a list to store batch requests
  batch_requests <- list()
  request_counter <- 1
  
  # Create n_repetitions requests for each essay (grouped by essay)
  for(i in 1:nrow(essays_df)){
    essay_prompt <- claude_all_rubric_prompt(essays_df$prompt[i], essays_df$essay[i])
    
    # Create 20 requests for this essay
    for(rep in 1:n_repetitions){
      request <- list(
        custom_id = paste0("essay_", essays_df$id[i], "_rep_", rep),
        params = list(
          model = "claude-sonnet-4-20250514",
          max_tokens = 1024,
          temperature = 0, # added low temp
          messages = list(
            list(
              role = "user",
              content = essay_prompt
            )
          )
        )
      )
      
      batch_requests[[request_counter]] <- request
      request_counter <- request_counter + 1
    }
  }
  
  cat("Created", length(batch_requests), "batch requests for", nrow(essays_df), "essays\n")
  
  return(batch_requests)
}

lowtemp_batch_requests <- create_lowtemp_requests(student_essays)
lowtemp_batch_jsonl <- write_batch_file(lowtemp_batch_requests)
lowtemp_batch_info <- submit_batch(lowtemp_batch_jsonl)
# Batch submitted successfully!
# Batch ID: msgbatch_0184nJ7Kui9v6WVVh47939mn 

lowtemp_status <- check_batch_status(lowtemp_batch_info$id)
lowtemp_start_time <- ymd_hms(lowtemp_status$created_at)
lowtemp_stop_time <- ymd_hms(lowtemp_status$ended_at)
lowtemp_run_time <- as.numeric(lowtemp_stop_time - lowtemp_start_time)
# Time difference of 1.473892 mins

lowtemp_batch_results <- get_batch_results(lowtemp_batch_info$id)
save(lowtemp_batch_results, file = "data/lowtemp_batch_results.Rdata")

lowtemp_batch_df <- extract_batch_scores(lowtemp_batch_results)
```
```{r, eval = FALSE, echo = FALSE}
save(lowtemp_batch_df, file = 'data/lowtemp_batch_df.Rdata')
```
```{r, echo = FALSE}
load('data/lowtemp_batch_df.Rdata')
```
```{r}
lowtemp_summary_detailed <- lowtemp_batch_df %>%
  group_by(essay_id) %>%
  summarise(
    n = n(),
    # Content
    content_mean = mean(content, na.rm = TRUE),
    content_sd = sd(content, na.rm = TRUE),
    content_min = min(content, na.rm = TRUE),
    content_max = max(content, na.rm = TRUE),
    # Organization
    org_mean = mean(organization, na.rm = TRUE),
    org_sd = sd(organization, na.rm = TRUE),
    org_min = min(organization, na.rm = TRUE),
    org_max = max(organization, na.rm = TRUE),
    # Language
    lang_mean = mean(language, na.rm = TRUE),
    lang_sd = sd(language, na.rm = TRUE),
    lang_min = min(language, na.rm = TRUE),
    lang_max = max(language, na.rm = TRUE),
    # Total
    total_mean = mean(total, na.rm = TRUE),
    total_sd = sd(total, na.rm = TRUE),
    total_min = min(total, na.rm = TRUE),
    total_max = max(total, na.rm = TRUE),
  ) %>%
  arrange(total_mean)

lowtemp_summary_detailed <- lowtemp_summary_detailed %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))

datatable(lowtemp_summary_detailed, 
          options = list(pageLength = 10, scrollX = TRUE),
          caption = "Summary Statistics by Essay (20 repetitions each) - Temperature = 0")

```

Now we're seeing more consistent scores, and it took _less than 90 seconds_ to process all 200 requests.

